NLTK: The Natural Language Toolkit

### 2004-05-25 technical/acl2004/acl04.tex
minor fixes in response to reviewers' comments and final proofread, plus formatting tweaks
svn/trunk@1822

\documentclass[11pt]{article}
\usepackage{acl04}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url,alltt,epsfig,boxedminipage}

@@

\date{}

@@

We chose Python because it has a shallow learning curve, its syntax
and semantics are transparent, and it has good string-handling
functionality.  As an interpreted language, Python facilitates
interactive exploration.  As an object-oriented language, data and
methods can be encapsulated and re-used easily.  Python comes with an extensive
standard library, including tools for graphical programming and
numerical processing.  And the recently added generator
syntax makes it easy to create interactive implementations of
algorithms \cite{Loper04,Rossum03intro,Rossum03ref}.

@@

To maximize interoperability between modules, we use a
single class to encode information about natural language texts -- the
\object{Token} class.  Each \object{Token} instance represents a
single unit of text, such as a word, sentence, or document; and is
defined by a (partial) mapping from property names to values.  For
example, the \object{TEXT} property is used to encode a token's text
content:\footnote{Some code samples are specific to NLTK
  version 1.4.}

@@

\begin{alltt}\small
\textbf{>>> from nltk.tokenizer import *}
\textbf{>>> token = Token(TEXT="Hello World!")}
\textbf{>>> WhitespaceTokenizer().tokenize(token)}
\textbf{>>> print token['SUBTOKENS'])}
[<Hello>, <World!>]
\end{alltt}
%
In a similar fashion, other language processing tasks such as
word-sense disambiguation, chunking and parsing all add properties to
the \object{Token} data structure.

In general, language processing tasks are formulated as
annotations and transformations involving \object{Tokens}.  In
particular, each processing task takes a token and extends it to
include new information.  These modifications are typically
\emph{monotonic}; in other words, new information is added but
existing information is not deleted or modified.  Thus, tokens serve
as a \emph{blackboard}, where information about a piece of text is
collated.  This architecture contrasts with the more typical
\emph{pipeline} architecture, where each processing task's output
discards its input information.  We chose the ``blackboard'' approach
over the ``pipeline'' approach because it allows more flexibility when
combining tasks into a single system.

@@

\begin{table*}
\small\noindent
\begin{boxedminipage}{\linewidth}
\begin{tabular}{llll}
\textbf{Corpus} &
\textbf{Contents and Wordcount} &
\textbf{Example Application} \\

@@

Brown Corpus &
15 genres, 1.15Mw, tagged &
training and testing taggers, text classification \\

CoNLL 2000 Chunking Data &
270kw, tagged and chunked &
training and testing chunk parsers \\

@@

NIST 1999 IEER (selection) &
63kw, named-entity markup &
training and testing named-entity recognizers \\

@@

Names Corpus &
8k male and female names &
text classification \\

@@

SEMCOR &
880kw, POS and sense tagged &
word-sense disambiguation \\

SENSEVAL 2 Corpus &
600kw, POS and sense tagged &
word-sense disambiguation \\

@@

Penn Treebank (sample) &
40kw, tagged and parsed &
parser development \\

Wordnet 1.7 &
180kw in a semantic network &
word-sense disambiguation, NL understanding \\

@@

Many language processing tasks must be developed and tested using
annotated data sets, or corpora.  Several such corpora are distributed
with NLTK, as listed in Table~\ref{tab:data}.  The
\object{corpus} module defines classes for reading and processing
many of these corpora.  The following code fragment illustrates
how the Brown Corpus is accessed.

@@

\begin{alltt}\small
\textbf{>>> from nltk.corpus import treebank}
\textbf{>>> treebank.groups()}
('raw', 'tagged', 'parsed', 'merged')
\textbf{>>> treebank.items('parsed')}
['wsj_0001.prd', 'wsj_0002.prd', ...]
\textbf{>>> item = 'parsed/wsj_0001.prd'}
\textbf{>>> sentences = treebank.tokenize(item)}
\textbf{>>> for sent in sentences['SUBTOKENS']:}
\textbf{...     print sent.pp()} \emph{# pretty-print}
(S:
  (NP-SBJ:
    (NP: <Pierre> <Vinken>)
    ...
\end{alltt}
%

@@

@@

NLTK has used at the University of Pennsylvania since 2001, and has
subsequently been adopted by several NLP courses at other
universities, including those listed in Table~\ref{tab:courses}.

Third party contributions to NLTK include: Brill tagger (Maloof),
hidden Markov model tagger (Cohn), GPSG-style feature-based grammar
and parser (Speer, Berwick), Kimmo finite-state morphological analyzer
(de Marcken, Yankama, Berwick) and decision list and decision tree
classifiers (Cohn).

NLTK is an open source project, and we welcome any contributions.
There are several ways to contribute: users can report bugs, suggest
features, or contribute patches on Sourceforge; users can participate
in discussions on the \textit{NLTK-Devel} mailing
list\footnote{\url{http://lists.sourceforge.net/lists/listinfo/nltk-devel}}
or in the NLTK public
forums; and users can submit their own NLTK-based projects for
inclusion in the nltk\_contrib directory.  New code modules that are
relevant, substantial, original and well-documented will be considered
for inclusion in NLTK proper.  Potential contributors should note that
all source code is distributed under the GNU General Public License,
and all documentation is distributed under a Creative Commons
non-commercial license.  Thus they can be confident that their
work will remain freely available to all.  Further information about
contributing to NLTK is available at \url{http://nltk.sf.net/contrib.html}.

\begin{table}[bt]
\small\noindent
\begin{boxedminipage}{\linewidth}
\begin{tabular}{l}
University of Pennsylvania, USA \\
\hspace{2ex}
\textit{Introduction to Computational Linguistics} \\[.5ex]

@@

@@

National Autonomous University of Mexico, Mexico \\
\hspace{2ex}
\textit{Introduction to Natural Language Processing in Python} \\[.5ex]

@@

NLTK is a broad-coverage toolkit that provides a simple, extensible,
uniform framework for assignments, demonstrations and projects.  It is
thoroughly documented, easy to learn, and simple to use.  Readers who
would like to receive occasional announcements about NLTK are
encouraged to sign up for the low-volume, moderated mailing list
\textit{NLTK-Announce}.\footnote{\url{http://lists.sourceforge.net/lists/listinfo/nltk-announce}}

@@

>>>>

\documentclass[11pt]{article}
\usepackage{colacl}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url,alltt,epsfig,boxedminipage}

@@

@@

We chose Python because it has a shallow learning curve, its syntax
and semantics are transparent, and it has good string-handling
functionality.  As an interpreted language, Python facilitates
interactive exploration.  As an object-oriented language, Python
permits data and methods to be encapsulated and re-used easily.  Python comes with an extensive
standard library, including tools for graphical programming and
numerical processing.  The recently added generator
syntax makes it easy to create interactive implementations of
algorithms \cite{Loper04,Rossum03intro,Rossum03ref}.

@@

To maximize interoperability between modules, we use a
single class to encode information about natural language texts -- the
\object{Token} class.  Each \object{Token} instance represents a
unit of text such as a word, sentence, or document, and is
defined by a (partial) mapping from property names to values.  For
example, the \object{TEXT} property is used to encode a token's text
content:\footnote{Some code samples are specific to NLTK
  version 1.4.}

@@

\begin{alltt}\small
\textbf{>>> from nltk.tokenizer import *}
\textbf{>>> tok = Token(TEXT="Hello World!")}
\textbf{>>> WhitespaceTokenizer().tokenize(tok)}
\textbf{>>> print tok['SUBTOKENS'])}
[<Hello>, <World!>]
\end{alltt}
%
In a similar fashion, other language processing tasks such as
word-sense disambiguation, chunking and parsing all add properties to
the \object{Token} data structure.

In general, language processing tasks are formulated as
annotations and transformations involving \object{Tokens}.  In
particular, each processing task takes a token and extends it to
include new information.  These modifications are typically
\emph{monotonic}; new information is added but
existing information is not deleted or modified.  Thus, tokens serve
as a \emph{blackboard}, where information about a piece of text is
collated.  This architecture contrasts with the more typical
\emph{pipeline} architecture where each processing task's output
discards its input information.  We chose the blackboard approach
over the pipeline approach because it allows more flexibility when
combining tasks into a single system.

@@

\begin{table*}
\small\noindent
\begin{boxedminipage}{\linewidth}
\begin{tabular}{llll}
\setlength{\tabcolsep}{0.5\tabcolsep}
\textbf{Corpus} &
\textbf{Contents and Wordcount} &
\textbf{Example Application} \\

@@

Brown Corpus &
15 genres, 1.15Mw, tagged &
training \& testing taggers, text classification \\

CoNLL 2000 Chunking Data &
270kw, tagged and chunked &
training \& testing chunk parsers \\

@@

NIST 1999 IEER (selection) &
63kw, named-entity markup &
training \& testing named-entity recognizers \\

@@

Names Corpus &
8k male \& female names &
text classification \\

@@

SEMCOR &
880kw, POS \& sense tagged &
word-sense disambiguation \\

SENSEVAL 2 Corpus &
600kw, POS \& sense tagged &
word-sense disambiguation \\

@@

Penn Treebank (sample) &
40kw, tagged \& parsed &
parser development \\

Wordnet 1.7 &
180kw in a semantic network &
WSD, NL understanding \\

@@

Many language processing tasks must be developed and tested using
annotated data sets or corpora.  Several such corpora are distributed
with NLTK, as listed in Table~\ref{tab:data}.  The
\object{corpus} module defines classes for reading and processing
many of these corpora.  The following code fragment illustrates
how the Brown Corpus is accessed.

@@

\begin{alltt}\small
\textbf{>>> from nltk.corpus import treebank}
\textbf{>>> treebank.groups()}
('raw', 'tagged', 'parsed', 'merged')
\textbf{>>> treebank.items('parsed')}
['wsj_0001.prd', 'wsj_0002.prd', ...]
\textbf{>>> item = 'parsed/wsj_0001.prd'}
\textbf{>>> sentences = treebank.tokenize(item)}
\textbf{>>> for sent in sentences['SUBTOKENS']:}
\textbf{...     print sent.pp()} \emph{# pretty-print}
(S:
  (NP-SBJ:
    (NP: <Pierre> <Vinken>)
    (ADJP:
      (NP: <61> <years>)
      <old>
    )
    ...
\end{alltt}
%

@@

NLTK includes the following modules:
\object{cfg},
\object{corpus},
\object{draw}
(\object{cfg},
\object{chart},
\object{corpus},
\object{featurestruct},
\object{fsa},
\object{graph},
\object{plot},
\object{rdparser},
\object{srparser},
\object{tree}),
\object{eval},
\object{featurestruct},
\object{parser}
(\object{chart},
\object{chunk},
\object{probabilistic}),
\object{probability},
\object{sense},
\object{set},
\object{stemmer}
(\object{porter}),
\object{tagger},
\object{test},
\object{token},
\object{tokenizer},
\object{tree}, and
\object{util}.
Please see the online documentation for details.

@@

NLTK has been used at the University of Pennsylvania since 2001, and has
subsequently been adopted by several NLP courses at other
universities, including those listed in Table~\ref{tab:courses}.

Third party contributions to NLTK include: Brill tagger (Chris Maloof),
hidden Markov model tagger (Trevor Cohn, Phil Blunsom), GPSG-style feature-based grammar
and parser (Rob Speer, Bob Berwick), finite-state morphological analyzer
(Carl de Marcken, Beracah Yankama, Bob Berwick), decision list and decision tree
classifiers (Trevor Cohn), and Discourse Representation Theory
implementation (Edward Ivanovic).

NLTK is an open source project, and we welcome any contributions.
There are several ways to contribute: users can report bugs, suggest
features, or contribute patches on Sourceforge; users can participate
in discussions on the \textit{NLTK-Devel} mailing
list\footnote{\url{http://lists.sourceforge.net/lists/listinfo/nltk-devel}}
or in the NLTK public
forums; and users can submit their own NLTK-based projects for
inclusion in the nltk\_contrib directory.  New code modules that are
relevant, substantial, original and well-documented will be considered
for inclusion in NLTK proper.
All source code is distributed under the GNU General Public License,
and all documentation is distributed under a Creative Commons
non-commercial license.  Thus, potential contributors can be confident that their
work will remain freely available to all.  Further information about
contributing to NLTK is available at \url{http://nltk.sf.net/contrib.html}.

\begin{table}[bt]
\small\noindent
\begin{boxedminipage}{\linewidth}
\begin{tabular}{l}

National Autonomous University of Mexico, Mexico \\
\hspace{2ex}
\textit{Introduction to Natural Language Processing}\\
\hspace{2ex}
\textit{in Python} \\[.5ex]

Ohio State University, USA \\
\hspace{2ex}
\textit{Statistical Natural Language Processing}\\[.5ex]

University of Malta, Malta \\
\hspace{2ex}
\textit{Natural Language Algorithms}\\[.5ex]

@@

University of Pennsylvania, USA \\
\hspace{2ex}
\textit{Introduction to Computational Linguistics} \\[.5ex]

@@

@@

NLTK is a broad-coverage natural language toolkit that provides a simple, extensible,
uniform framework for assignments, demonstrations and projects.  It is
thoroughly documented, easy to learn, and simple to use.  NLTK is now widely
used in research and teaching.  Readers who
would like to receive occasional announcements about NLTK are
encouraged to sign up for the low-volume, moderated mailing list
\textit{NLTK-Announce}.\footnote{\url{http://lists.sourceforge.net/lists/listinfo/nltk-announce}}

@@

\pagebreak
